---
title: "The Value of Information"
subtitle: "for Conservation & Natural Resource Management"
author: "Will Morris"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "voi-seminar.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```
class: inverse, center, middle

# Part 1: Value of information
## A little background & theory

???


---
# Formal VOI

--

## Expected Value of Information (EVI)

--

$\Large\mathrm{EVI} = \mathrm{EVWNI} - \mathrm{EVWOI}$

--

## Expected Value With Original Information (EVWOI)

--

$\Large\mathrm{EVWOI} = \max_a\mathrm{E}_s[v(a,s)]$

???


---
# Formal VOI

--

## Many Forms of VOI

--

$\Large\mathrm{EV?I} = \mathrm{EVW?I} - \mathrm{EVWOI}$

--

## Expected Value of Perfect Information (EVPI)

--

$\Large\mathrm{EVPI} = \mathrm{EVWPI} - \mathrm{EVWOI}$

--

$\Large\mathrm{EVPI} = \mathrm{E}_s\max_a[v(a,s)] - \max_a\mathrm{E}_s[v(a,s)]$

???



---
# An example 

--
- .font130[Manager responsible for endangered species]

--
- .font130[Currently 10 breeding pairs, but declining]

--
- .font130[Must choose whether or not to restore habitat]

--
- .font130[50% chance that restoration will be a success]

--
- .font130[Expected number of breeding pairs:]

</br>

.font110[

| Action     | Failure | Success |
|:-----------|:-------:|:-------:|
| Do nothing |    8    |    8    |
| Restore    |    5    |    15   |
]

???



---
count: false
# An example

.font130[
- Manager responsible for endangered species
- Currently 10 breeding pairs, but declining
- Must choose whether or not to restore habitat
- 50% chance that restoration will be a success
- Expected number of breeding pairs:
]
</br>
.font110[

| Action     | Failure | Success | _EV_ |
|:-----------|:-------:|:-------:|:----:|
| Do nothing |    8    |    8    | _8_  |
| Restore    |    5    |    15   | _10_ |
]

--

$\Large\mathrm{EVWOI} = \max_a\mathrm{E}_s[v(a,s)]$

???



---
count: false
# An example

.font130[
- Manager responsible for endangered species
- Currently 10 breeding pairs, but declining
- Must choose whether or not to restore habitat
- 50% chance that restoration will be a success
- Expected number of breeding pairs:
]
</br>
.font110[

| Action     | Failure | Success | _EV_ |
|:-----------|:-------:|:-------:|:----:|
| Do nothing |    8    |    8    | .red[_8_]  |
| Restore    |    5    |    15   | .red[_10_] |
]

$\Large\mathrm{EVWOI} =\max_a$ .font150[(.red[8], .red[10])]

???



---
count: false
# An example

.font130[
- Manager responsible for endangered species
- Currently 10 breeding pairs, but declining
- Must choose whether or not to restore habitat
- 50% chance that restoration will be a success
- Expected number of breeding pairs:
]
</br>
.font110[

| Action     | Failure | Success | _EV_ |
|:-----------|:-------:|:-------:|:----:|
| Do nothing |    8    |    8    | _8_  |
| Restore    |    5    |    15   | .red[_10_] |
]

$\Large\mathrm{EVWOI} =$ .font150.red[10]

???



---
count: false
# An example

.font130[
- Manager responsible for endangered species
- Currently 10 breeding pairs, but declining
- Must choose whether or not to restore habitat
- 50% chance that restoration will be a success
- Expected number of breeding pairs:
]
</br>
.font110[

| Action     | Failure | Success | _EV_ |
|:-----------|:-------:|:-------:|:----:|
| Do nothing |    8    |    8    | _8_  |
| Restore    |    5    |    15   | .red[_10_] |
]

$\Large\mathrm{EVPI} = \mathrm{EVWPI} -$ .font150.red[10]

???



---
count: false
# An example

.font130[
- Manager responsible for endangered species
- Currently 10 breeding pairs, but declining
- Must choose whether or not to restore habitat
- 50% chance that restoration will be a success
- Expected number of breeding pairs:
]
</br>
.font110[

| Action     | Failure | Success | _EV_ |
|:-----------|:-------:|:-------:|:----:|
| Do nothing |    8    |    8    | _8_  |
| Restore    |    5    |    15   | .red[_10_] |
]

$\Large\mathrm{EVPI} = \mathrm{E}_s\max_a[v(a,s)] -$ .font150.red[10]

???



---
count: false
# An example

.font130[
- Manager responsible for endangered species
- Currently 10 breeding pairs, but declining
- Must choose whether or not to restore habitat
- 50% chance that restoration will be a success
- Expected number of breeding pairs:
]
</br>
.font110[

| Action     | Failure | Success | _EV_ |
|:-----------|:-------:|:-------:|:----:|
| Do nothing |    .blue[8]    |     8   | _8_  |
| Restore    |    5    |    .blue[15]   | .red[_10_] |
]

$\Large\mathrm{EVPI} = \mathrm{E}_s$.font150[(.blue[8], .blue[15]) - .red[10]]

???



---
count: false
# An example

.font130[
- Manager responsible for endangered species
- Currently 10 breeding pairs, but declining
- Must choose whether or not to restore habitat
- 50% chance that restoration will be a success
- Expected number of breeding pairs:
]
</br>
.font110[

| Action     | Failure | Success | _EV_ |
|:-----------|:-------:|:-------:|:----:|
| Do nothing |    .blue[8]    |     8   | _8_  |
| Restore    |    5    |    .blue[15]   | .red[_10_] |
]

$\Large\mathrm{EVPI} =$ .font150[.blue[11.5] - .red[10]]

???



---
count: false
# An example

.font130[
- Manager responsible for endangered species
- Currently 10 breeding pairs, but declining
- Must choose whether or not to restore habitat
- 50% chance that restoration will be a success
- Expected number of breeding pairs:
]
</br>
.font110[

| Action     | Failure | Success | _EV_ |
|:-----------|:-------:|:-------:|:----:|
| Do nothing |    .blue[8]    |     8   | _8_  |
| Restore    |    5    |    .blue[15]   | .red[_10_] |
]

$\Large\mathrm{EVPI} =$ .font150.purple[1.5]

???



---
# Formal VOI

--

## Expected Value of Sample Information

--

$\Large\mathrm{EVSI} = \mathrm{EVWSI} - \mathrm{EVWOI}$

--

$\Large\mathrm{EVSI} = \mathrm{E}_x[\max_a\mathrm{E}_{s|x}[v(a,s)]] - \max_a\mathrm{E}_s[v(a,s)]$


???

---
class: centre, middle

```{r dimretplot, fig.width = 4, fig.height = 3, dev = "svg", echo = FALSE, out.width = "1200px", dev.args = list(family = "Latin Modern Roman")}
par(mar = c(3, 3, 1, 1))
curve(x / (x + 1), 0, 10, xlab = "", xaxt = 'n', ylab = "", yaxt = "n", ylim = 0:1, lwd = 2)
abline(h = 1, lty = 2)
text(0.5, 1, "EVPI", pos = 1)
mtext("Sample-size", 1, 1)
mtext("EVSI", 2, 1)
```

---
class: inverse, center, middle

# Part 2: Value of Information in conservation and natural resource management
## How has it been used to date?

???



---
# Two flavours of VOI in conservation science

--

.pull-left[
## 1. Informal VOI analyses

- Don't use classic VOI (EVPI, etc.,)
- Post-hoc reevaluation of decision problems with different data
- Typically applied to spatial conservation planning (e.g., reserve network design)
- Somewhat hidden in literature as dont use a standard lexicon
- Often report a large value of information
]
--

.pull-right[
## 2. Formal VOI Analyses 

- Use one or more forms of EV?I in the strict sense
- Almost always calculate the EVPI
- Often applied to problems of managing a species (e.g., fishery, invasive plant)
- Easier to find in literature
- Typically report a low value of information
]

---
class: inverse, center, middle

# Part 3: A Case-study
## Value of Information and spatial conservation planning

???

---


---
class: inverse, center, middle

# Part 4: Value of Information and risk
## A counter-intuitive example 


???


